We describe here the predicates used to represent the physical network and how virtual networks are embedded on it.

For an eased reading for the rest of this section, we define some variable names used later on:

\begin{lstlisting}[backgroundcolor = \color{lightgray}]
vnet is the unique identifier of a virtual network.
pnode_name is the unique identifier of a physical node.
plink_name is the unique identifier of a physical link.
vnode_name is the unique identifier of a virtual node.
vlink_name is the unique identifier of a virtual link.
ti is the unique identifier of a time interval.
\end{lstlisting}

\paragraph{Physical network modeling}


The physical network is composed of nodes and links. We define two temporal predicates:

\begin{lstlisting}[backgroundcolor = \color{lightgray}]
physical_node(node_name,ti)
physical_link(plink_name,node_name1,node_name2,ti)
\end{lstlisting}

These predicates represent the existence of each element and failures or attacks will limit this existence.

For instance, if node1 is attacked at $t_1$ and only becomes available again at $t_2$ until the next failure/attack at $t_3$ then we obtain the following predicates:

\begin{lstlisting}[backgroundcolor = \color{lightgray}]
physical_node(node1,[0,t_1])
physical_node(node1,[t_2,t_3])
\end{lstlisting}

$t_3$ could be seen as $+\infty$ if no other incident occurs on node1.

\paragraph{Virtual network modeling}
We extend the previous predicates with information related to the tenant of the virtual network and it is embedded on top of a set of physical resources.


\begin{lstlisting}[backgroundcolor = \color{lightgray}]
virtual_network(vnet_name,ti)
virtual_node(vnode_name,vnet_name,ti)
virtual_link(vlink_name,vnode_name1,vnode_name2,vnet_name,ti)
\end{lstlisting}

\paragraph{Embedding modeling}
Each virtual element, node or link, is embedded on a set of physical resources.
Therefore, there can be one or more predicates used to describe the embedding of a virtual element.

\begin{lstlisting}[backgroundcolor = \color{lightgray}]
node_embedding(vnode_name,pnode_name,ti)
node_embedding(vnode_name,plink_name,ti)
link_embedding(vlink_name,pnode_name,ti)
link_embedding(vlink_name,plink_name,ti)
\end{lstlisting}

\paragraph{Example for modeling physical and virtual networks}

We consider the use case of Fig.~\ref{fig:VNE-example-model}.
There is no failure in the system nor any attack, therefore, we consider the time interval ti to represent the duration of the predicates.

\input{figures/vne_example_model.tex}

The current infrastructure gives the following output:

\begin{lstlisting}[backgroundcolor = \color{lightgray}]
virtual_network(vSDN1,ti)
virtual_network(vSDN2,ti)
virtual_node(V1,vSDN1,ti)
virtual_node(V2,vSDN1,ti)
virtual_node(V3,vSDN1,ti)
virtual_node(V4,vSDN2,ti)
virtual_node(V5,vSDN2,ti)
virtual_node(V6,vSDN2,ti)
virtual_link(vlink1,V1,V2,vSDN1,ti)
virtual_link(vlink2,V1,V3,vSDN1,ti)
virtual_link(vlink3,V4,V5,vSDN2,ti)
virtual_link(vlink4,V5,V6,vSDN2,ti)
physical_node(P1,ti)
physical_node(P2,ti)
physical_node(P3,ti)
physical_node(P4,ti)
physical_node(P5,ti)
physical_node(P6,ti)
physical_link(plink1,P1,P2,ti)
physical_link(plink2,P1,P3,ti)
physical_link(plink3,P2,P4,ti)
physical_link(plink4,P3,P4,ti)
physical_link(plink5,P4,P5,ti)
physical_link(plink6,P4,P6,ti)
node_embedding(V1,P1,ti)
node_embedding(V4,P1,ti)
node_embedding(V2,P2,ti)
node_embedding(V3,P3,ti)
node_embedding(V5,P4,ti)
node_embedding(V6,P5,ti)
link_embedding(vlink1,plink1,ti)
link_embedding(vlink2,plink2,ti)
link_embedding(vlink3,plink1,ti)
link_embedding(vlink3,plink4,ti)
link_embedding(vlink4,plink5,ti)
\end{lstlisting}