\documentclass[a4paper, 11pt]{report}
\usepackage{comment} % enables the use of multi-line comments (\ifx \fi) 
% \usepackage{lipsum} %This package just generates Lorem Ipsum filler text. 
\usepackage{fullpage} % changes the margin
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{url}
\usepackage[toc,page]{appendix}
\usepackage{algpseudocode}
\usepackage[ruled,noend,linesnumbered]{algorithm2e}
\usepackage{listings}
\usepackage{lastpage}
\usepackage{fancyhdr}
\usepackage[Lenny]{fncychap}
\usepackage[most]{tcolorbox}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{color}
\usepackage{xspace}
\usepackage{parskip}
\usepackage{afterpage}
% \usepackage{refcount}
\usepackage{paralist}
\usepackage{authblk}
\usepackage{etoolbox}
\usepackage{listings}
\usepackage{lipsum}
\usepackage{enumitem}
%\usepackage{fancyhdr}% http://ctan.org/pkg/fancyhdr
\usepackage{verbatim}% http://ctan.org/pkg/verbatim
\usepackage{multirow}
% \usepackage{slashbox}
\usepackage{caption}
\usepackage{cite}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.15} 
\usepgfplotslibrary{patchplots}
% Extra depth for titles
\usepackage[compact]{titlesec}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{2}
\usepackage[nottoc]{tocbibind}
% \usepackage[style=alphabetic,citestyle=authoryear,doi=false]{biblatex}
\usepackage{minted}
\usepackage{epigraph}
\usepackage{quotchap}
\usepackage{setspace}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true, %set true if you want colored links
    linktoc=all,     %set to all if you want both sections and subsections linked
    linkcolor=black, %choose some color if you want links to stand out
}
\usepackage{cleveref}

\usepackage{etoolbox} % provides \patchcmd
\patchcmd{\endabstract}{\null}{}{}{} % removing blank page after abstract

\lstset{numbers=left,numberstyle=\tiny,numbersep=5pt,language=Lisp,
  stringstyle=\ttfamily\small,basicstyle=\ttfamily\footnotesize,
  showstringspaces=false,breaklines=true,keywordstyle=\color{black},}

\setstretch{1.21}
%  MARGIN
\addtolength{\textwidth}{-0.4in}

\newtcolorbox{myformula}{colback=block-gray,boxrule=0pt,boxsep=0pt,breakable,halign= flush left,before skip=10pt, after skip=20pt}
\definecolor{block-gray}{gray}{0.95}

\DeclareMathOperator*{\argmax}{arg\,max}

\newcommand{\boxedtext}[1]{\fbox{\scriptsize\bfseries\textsf{#1}}}
\newcommand{\myremark}[2]{
   \textcolor{blue}{\boxedtext{#1}
      {\small$\blacktriangleright$\emph{\textsl{#2}}$\blacktriangleleft$}
}}
\newcommand{\DELETE}[2]{
   \textcolor{red}{
         {\small$\blacktriangleright$\st{#2}$\blacktriangleleft$}
    }
}
\newcommand\GB[1]{\myremark{GB}{#1}}
\newcommand\CK[1]{\myremark{CK}{#1}}
\newcommand\FC[1]{\myremark{FC}{#1}}
% \newcommand\RW[1]{\myremark{RW}{#1}}

\newcommand{\etc}{etc.\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etal}{et al.\xspace}
\newcommand{\eg}{e.g.,\xspace}

\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

% \renewenvironment{equation}{\color{black}}{}
\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem*{definition}{Definition}


\title{\LARGE \bf Security characterization of SDN Virtual Network migration:\\ Formal Approach and Resource Optimization}

\author{Fabien Charmet}

% \addbibresource{references.bib} %Imports bibliography file
\begin{document}

% \input{frontcover.tex}

%Header-Make sure you update this information!!!!
% \noindent
\maketitle
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
\pagenumbering{roman}
This thesis investigates the security of Virtual Network migration. Over the years, virtualization has been used to optimize physical resources and to support businesses’ infrastructure. Virtualization consists in exposing a fraction of a resource for a user to operate. Virtual Machines are used to host business services like web servers or a backup service. Network virtualization has not benefited from the same interest from researchers and the industry. The Software Defined Networking paradigm has introduced new possibilities to implement network virtualization and provide users with a flexible network, decoupled from the physical equipment. Virtual Networks are used to interconnect Virtual Machines and can be configured with specific routing policies or security protocols. In case of a failure of the resource, either accidental or intentional, the virtualization infrastructure will migrate the resource to maintain the service provided.

The security of Virtual Machines and their migration is a well-researched topic that has been widely in the past, while the study of network virtualization and especially the migration process only are at an early stage. The attack surface of network virtualization is similar in nature to the virtualization of legacy resources, and presents an additional aspect because of the use of Software Defined Networking.
The motivation of this research is to investigate the security of the Virtual Network migration process in the context of Software Defined Networking. In order to do so, we first define the scope of the study and focus on the networking aspect of the migration. Then, we outline the threat model of the migration process and devise a detection mechanism against attacks in the virtualization infrastructure. Finally, we optimize the previous mechanism by optimizing the deployment of network monitoring resources for an optimal coverage.
    
In the first part of this thesis we propose a formal approach to describe the different aspects of the virtualization infrastructure. We use a first order formalism to model several security properties as a set of logical predicates. These predicates account for both physical and virtual elements of the virtualization infrastructure, and the data use by both end users and the infrastructure owner.
An execution trace is generated during the migration of a Virtual Network, and will be used by a theorem prover to compute a formal proof to verify if a security violation occurred. The first order model is based on the assumption that the execution trace is generated using perfect monitoring. This implies that the proof is complete and that the networking monitoring is always done under optimal conditions.
We alleviate this assumption by modeling a resource allocation problem to determine how the monitoring resources should be deployed and which network nodes provide the best coverage. We solve this problem using a Markov Decision Process, and determine a dynamic deployment of monitoring resources during the migration. We conclude our optimization with a proposition of a static deployment of the resources prior to the migration.
% We propose to model the virtualization infrastructure and describe the security of the migration process using a first order temporal logic. We then use this logic with a theorem prover to verify if a security violation occurred during the migration of a network.  We use a Markov Decision Process to determine the optimal placement of monitoring resources in the infrastructure.

\setcounter{page}{2}
\begin{savequote}
--Do not worry about your difficulties in mathematics, I can assure you that mine are all greater.
\qauthor{Albert Einstein}
\end{savequote}
\chapter*{R\'esum\'e}
\addcontentsline{toc}{chapter}{R\'esum\'e}
Cette th\`{e}se explore la s\'{e}curit\'{e} de la migration de r\'{e}seaux virtuels. Au cours des ann\'{e}es, la virtualisation a \'{e}t\'{e} utilis\'{e}e pour optimiser l'usage des ressources informatiques et pour supporter les infrastructures des entreprises. La virtualisation consiste \`{a} allouer une partie des ressources d'une machine physique \`{a} un utilisateur (sous la forme d'une machine virtuelle) pour qu'il puisse l'exploiter. Les machines virtuelles sont utilis\'{e}es pour h\'{e}berger des services op\'{e}rationnels comme un serveur internet ou une base de donn\'{e}es. La virtualisation des r\'{e}seaux n'a pas profit\'{e} du m\^{e}me int\'{e}r\^{e}t de la part des chercheurs et des acteurs industriels. Le paradigme du Software Defined Networking a introduit de nouvelles possibilit\'{e}s pour impl\'{e}menter la virtualisation r\'{e}seau et fournir aux utilisateurs une solution flexible pour leurs besoins m\'{e}tiers. Les r\'{e}seaux virtuels sont utilis\'{e}s pour interconnecter des machines virtuelles, et ils peuvent \^{e}tre configur\'{e}s avec des r\`{e}gles de routages ou des protocoles de s\'{e}curit\'{e} sp\'{e}cifiques. Dans l'\'{e}ventualit\'{e} o\`{u} un \'{e}quipement r\'{e}seau tomberait en panne ou sous le coup d'une attaque informatique, le syst\`{e}me d'hypervision va migrer les ressources afin de pr\'{e}server la disponibilit\'{e} des services utilisateurs.

La s\'{e}curit\'{e} des machines virtuelles et de leur migration est un domaine de recherche qui a \'{e}t\'{e} grandement explor\'{e} par le pass\'{e}, tandis que la virtualisation r\'{e}seau et plus sp\'{e}cifiquement la migration de r\'{e}seaux virtuels restent encore des domaines de recherche assez jeune et o\`{u} beaucoup reste \`{a} faire. La surface d'attaque de la virtualisation r\'{e}seau est similaire en nature \`{a} celle de la virtualisation traditionnelle, mais elle pr\'{e}sente un aspect suppl\'{e}mentaire d\^{u} \`{a} l'usage du paradigme du Software Defined Networking. La motivation de notre travail est d'\'{e}tudier la s\'{e}curit\'{e} du processus de migration des r\'{e}seaux virtuels, dans le contexte du Software Defined Networking. Nous proposons d'atteindre cet objectif en trois phases. Tout d'abord, nous d\'{e}finissons le p\'{e}rim\`{e}tre de cette \'{e}tude, et nous concentrons sur l'aspect r\'{e}seau de la migration. Ensuite, nous d\'{e}crivons le mod\`{e}le d'attaquant afin de compromettre la migration des r\'{e}seaux et nous concevons un m\'{e}canisme de d\'{e}tection contre les attaques envers l'infrastructure de virtualisation. Enfin, nous am\'{e}liorons le m\'{e}canisme de d\'{e}fense en optimisant le d\'{e}ploiement des ressources de d\'{e}tection afin d'obtenir une couverture optimale de l'infrastructure. 

Dans la premi\`{e}re partie de cette th\`{e}se, nous proposons une approche formelle pour d\'{e}crire les diff\'{e}rents composants de l'infrastructure de virtualisation.
Nous utilisons un formalisme de logique du premier ordre pour d\'{e}crire diff\'{e}rentes propri\'{e}t\'{e}s de s\'{e}curit\'{e} sous la forme de pr\'{e}dicats bool\'{e}ens. 
Cette mod\'{e}lisation inclut la repr\'{e}sentation des donn\'{e}es des utilisateurs finaux ainsi que l'infrastructure de virtualisation.
Une trace d'ex\'{e}cution est g\'{e}n\'{e}r\'{e}e pendant la migration d'un r\'{e}seau virtuel, et est ensuite utilis\'{e}e par un prouveur de th\'{e}or\`{e}me afin de v\'{e}rifier formellement si la s\'{e}curit\'{e} de la migration a \'{e}t\'{e} respect\'{e}e. Le mod\`{e}le formel est bas\'{e} sur la supposition que la trace d'ex\'{e}cution est g\'{e}n\'{e}r\'{e}e par un outil de supervision exempt de tout d\'{e}faut. Ceci implique que la preuve formelle est compl\`{e}te et que la supervision r\'{e}seau est toujours r\'{e}alis\'{e}e dans des conditions optimales. Nous levons cette hypoth\`{e}se en mod\'{e}lisant un probl\`{e}me d'allocation de ressources afin de d\'{e}terminer quels \'{e}quipements r\'{e}seau devraient \^{e}tre charg\'{e}s de la d\'{e}tection d'attaques pour une couverture optimale. Nous r\'{e}solvons ce probl\`{e}me en utilisant un processus de d\'{e}cision markovien et d\'{e}terminons un d\'{e}ploiement dynamique des ressources pendant la migration. Nous concluons notre optimisation en proposant un d\'{e}ploiement statique des ressources, en amont de toute migration.






\chapter*{}
\setcounter{page}{4}
\addcontentsline{toc}{chapter}{Dedication}
\thispagestyle{empty}

\begin{center}
\textbf{\\[2in]}
\textit{I dedicate this work to my father, whom I miss very much.}    
\end{center}

\begin{savequote}
--Showing gratitude is one of the simplest yet most powerful things humans can do for each other.
\qauthor{Randy Pausch}
\end{savequote}
\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}


As I finally reach the end of this long yet interesting journey, I realize I have so many people to thank.

I would like to thank my PhD supervisor, \textbf{Prof. Abdallah M'Hamed}, for his wisdom and his advice that made every interaction interesting.

I also deeply thank \textbf{Dr. Gregory Blanc} and \textbf{Dr. Christophe Kiennert}, without whom this work 
would never have been possible. Their remarks, their patience and their tremendous support helped me throughout this long path. They made me realize what matters during the PhD life and what it means to set high expectations for myself. There are no words for the gratitude I have towards them.

I thank \textbf{Prof. Viet Triem Tong} and \textbf{Prof. Meddahi} who reviewed my thesis and provided me with valuable remarks to improve it, and thanks to \textbf{Prof. Chrisment}, \textbf{Prof. Neves} and \textbf{Dr. Leneutre} for attending my defence.

This whole opportunity would not have been possible without the trust of \textbf{Prof. Herv\'e Debar}, who gave me a chance to become an Engineer and then a Doctor. I have had premium hands on experience on research and infrastructure deployment. I owe him all these years spent at T\'el\'ecom SudParis.

I am grateful for the collaboration I had with \textbf{Dr. Richard Waldinger} at the Stanford Research Institute. His expertise in theorem proving and his remarks made this work possible. I enjoyed every moment of our collaboration.

I have so many colleagues to thank here, for the support and cheerfulness they brought in along the way.
Thank you \textbf{Nesrine}, \textbf{Khalifa}, \textbf{Gamain}, \textbf{Miroslav}, \textbf{Stef}, \textbf{Seb}, \textbf{Christian}, \textbf{Olivier} and so many others. I also wish good luck to \textbf{Mariana}, \textbf{Antoine}, \textbf{Mus}, \textbf{Maria}, \textbf{Hadjer}, \textbf{Aina}, \textbf{Souha}, who should cross this line one day or another.

A big thank to all of my friends, in no particular order: \textbf{Guillaume}, \textbf{Jo}, \textbf{Thomas}, \textbf{Nico}, \textbf{Yoan}, \textbf{Chan}, \textbf{Etienne}, whose company made me laugh when I needed it the most. 

I am grateful towards \textbf{Vanessa Sanfilippo}, who lifted me up through music and who has been an extraordinary piano teacher. I only wish to meet her expectations towards me.

A very special mention to my family, which went through thick and thin with me, and to whom my love is unconditional. 

Finally, there are three persons I have to thank for their incredible work during my PhD: \textbf{Fran\c{c}oise Abad}, \textbf{Sandra Gchweinder} and \textbf{Zahia Zendagui}. They made every single day run smoothly.

All mistakes are mine, enjoy your reading !

\newpage

\tableofcontents
\listoffigures
 \listoftables
\thispagestyle{empty}
% \pagestyle{empty}


\newpage
\pagenumbering{arabic}
\begin{savequote}
--Every action needs to be prompted by a motive.
\qauthor{Leonardo da Vinci}
\end{savequote}
\chapter{Introduction}
\input{introduction.tex}
 

\newpage
\begin{savequote}[0.5\textwidth]
--A man cannot understand the art he is studying if he only looks for the end result without taking the time to delve deeply into the reasoning of the study.
\qauthor{Miyamoto Musashi}
\end{savequote}
\chapter{State of the Art}
\label{sec:sota}
In this section, we provide a literature review of the different topics studied in this thesis.
First of all, we start by defining the different networking and virtualization concepts considered in this thesis.
Secondly, we examine the existing network hypervisors, and detail them under the scope of network migration with regard to the type of abstraction they provide.
Thirdly, we consider works that investigate the Virtual Network migration process, with regard to the purpose of the migration, whether it is for performance optimization, technical limitations evaluation or for a formal study of the process.
Fourthly, we investigate the security of the Virtual Network migration, namely by considering the attack surface of a similar problem, Live Virtual Machine migration, as well as the formal models used to express security properties in a computer network.
Both topics outline the needs for a security study of the VN migration problem.
The last topic considered here is related to the resource allocation problem in a network infrastructure, and we present different formalisms used in this context.
This problem is studied to improve the applicability of a model\GB{what model?} to a real life environment.

\section{Definitions}
\label{sec:basic_def}
\input{definitions.tex}

\section{SDN-based Network Hypervisors}
% \GB{I am still incomfortable with the fact that this is the same section title as \ref{def:netvirt}. The section titles need to be distinct as their contents are distinct.}

Network virtualization has leveraged the use of SDN to provide a flexible abstraction layer to service providers. In this section, we first propose classification criteria for network hypervisors and then evaluate existing hypervisors based on this classification. We will then propose a reference architecture for network hypervisors implementing a secure network migration feature, based on our survey of existing solutions.

\input{sota/sota-netvirt-sdn/introduction.tex}

\input{sota/sota-netvirt-sdn/existing_nhv.tex}

\input{sota/sota-netvirt-sdn/existingnvh-table.tex}

\input{sota/sota-netvirt-sdn/nhv_refarchi.tex}

\input{sota/sota-netvirt-sdn/summary_refarchi.tex}

\input{sota/sota-netvirt-sdn/refarchi_table.tex}

\input{sota/sota-netvirt-sdn/existing_vnmigration.tex}

\section{Network \& VM security}

We have presented, in Sections~\ref{sec:existing-nhv} and~\ref{sec:sota-vnmigration}, existing solutions for network virtualization and Virtual Network migration. We now delve into the security aspect of network virtualization, and the related security topics. The security of VN migration is a topic that has not been studied yet, while the security of VM migration has been thoroughly investigated. The following sections cover two topics, namely the security of the Live VM migration process and existing formal models for network security.
Existing formal models are useful to understand to which extent security in networks has been expressed but also it will help us to determine what has not beed modeled yet\GB{maybe split the last sentence in two or more sentences because it is confusing, as is.}.

\subsection{Security of VM migration}

\input{sota/sota-VM/attacking_vmmigration.tex}

\subsection{Formal Models for network security}

\input{sota/sota-FM/existing_formalmodels.tex}

\section{Resource Allocation in networking infrastructures}
The resource allocation (RA) problem consists in determining the optimal way to spend limited resources in order to achieve maximum profit while satisfying a set of constraints.
The VNE problem is a subclass of the RA problem\GB{sorry I suggested this sentence, but the RA you wish to use is not about the VNE}, where the VNE component will determine the set of physical resources that will embed a Virtual Network, while maximizing the acceptance rate, \ie the ratio between accepted VN requests and total number of requests.
The VNE problem is constrained by available resources, but there might be other constraints like security requirements or geographical positioning rules.
In this section, we consider two different formalisms used to solve the RA problem, namely Game Theory and Markov Decision Processes. 

% \input{sota/sota-RA/LP.tex}
\input{sota/sota-RA/GT.tex}
\input{sota/sota-RA/MDP.tex}

\begin{savequote}

--Formalism is music that people don’t understand at first hearing.
\qauthor{Boris Schwarz}
\end{savequote}
\section{General Conclusion}
\input{sota/sota-conclusion.tex}

\newpage
\chapter{Formal modeling of the migration process and its security}
\label{sec:formal_model}
\input{model/model-introduction.tex}

% \section{Methodology}
% \input{model/model-methodology.tex}

\section{Assumptions}
\input{model/model-assumptions.tex}

\newpage
\section{Modeling physical and Virtual Networks}
\input{model/model-networks.tex}

\newpage
\section{Modeling the migration of Virtual Networks}
\input{model/model-migration.tex}

\newpage
\section{Security properties for Virtual Network migration}
\input{model/model-security-properties.tex}

\input{figures/virtualization-attack-surface.tex}

\section{Threat model against the migration}
\input{model/model-attack-model.tex}

\section{Problem verification using a theorem prover}
\input{model/model-snark.tex}

\section{Use cases: Detection of security violations}
\input{model/model-usecase.tex}

\newpage
\section{General Conclusion}
\input{model/model-conclusion.tex}

\newpage
\begin{savequote}
--Premature optimization is the root of all evil.”
\qauthor{Donald Ervin Knuth}
\end{savequote}
\chapter{Optimizing monitoring resource allocation}
\label{sec:RAprob}
\input{RA/RA-motivation.tex}

\input{RA/GT.tex}

\newpage
\input{RA/MDP.tex}

\newpage
\begin{savequote}
--How to play music may be known. At the commencement of the piece, all the parts should sound together. As it proceeds, they should be in harmony while severally distinct and flowing without break, and thus on to the conclusion.
\qauthor{Confucius}
\end{savequote}
\chapter{Conclusion}
\input{conclusion.tex}

\newpage
\chapter*{List of publications}
\addcontentsline{toc}{chapter}{List of Publications}
\textbf{Fabien Charmet}, Gregory Blanc, Christophe Kiennert (2019). Optimizing Resource Allocation for Secure SDN-based Virtual Network Migration. \textit{IEEE 18th International Symposium on Network Computing and Applications}\\

\textbf{Fabien Charmet}, Richard Waldinger, Gregory Blanc, Christophe Kiennert, Khalifa Toumi (2017). Preserving confidentiality during the migration of virtual SDN topologies: A formal approach. \textit{IEEE 16th International Symposium on Network Computing and Applications}

\newpage
\input{appendixes/glossary.tex}

\newpage
% \FC{Don't forget to regenerate bibfile and change doi to optidoi}
\bibliography{test}{}
\bibliographystyle{alphaurl}


% \printbibliography

% \newpage
% \begin{appendices}
% \input{appendixes/glossary.tex}
% \newpage
% % \input{appendixes/snark-code.tex}
% \newpage
% \input{appendixes/notations.tex}
% \end{appendices}


\end{document}
\textbf{\\[1.5in]}

