\section{MDP notations summary}
\begin{itemize}
    \item $n$: The number of nodes in the infrastructure.
    \item $\textbf{N} = \{1,..,n\}$ is the set of nodes in the infrastructure.
    \item $S$ is the finite set of states of the system. 
    \item $Mi^s \subset \textbf{N} $ is the set of currently migrated nodes for state $s\in S$.
    \item $Mo^s \subset \textbf{N}$ is the set of currently monitored nodes for state $s\in S$.
    \item $At^s \subset \textbf{N}$ is the set of compromised nodes for state $s \in S$.
    \item $A = \{m_1,..,m_n,u_1,..,u_n,d\}$: is the set of actions available at each state, where $m_i$ is monitoring node $i$, $u_i$ is unmonitoring node $i$ and $d$ is do nothing.
    \item $P(s,s',a)$ is the transition probability from $s$ to $s'$ for action $a$.
    \item $R(s,a,s')$ is the reward function for action $a$ from state $s$ to $s'$.
    \item $s=<b_f,b_c,Mi,Mo,At>$ is a tuple describing the state of the system, $s\in S$.
    %\item $Lu$ is the set of currently lured nodes in the infrastructure
    \item $b_f$ is the financial budget, $b_f^s$ is the remaining financial budget of state $s$.
    % \GB{I don't understand the latter notation. Is it a typo? Did you mean $s_{b_f}$?}.
    \item $b_c$ is the computational power, $b_c^s$ is the remaining computational power of state $s$.
    % \GB{again, did you mean $s_{b_c}$?}.
    \item $c_f^j$, $c_c^j$ and $c_d$ are the costs impacted respectively on $b_f$ and $b_c$ to monitor/unmonitor node $j$ or do nothing.\FC{$c_f >> c_c$}
    \item $k_{fb}$ and $k_{cb}$ are the atomic costs proportional to respectively $c_f^j$ and $c_c^j$.
    % \GB{what does it mean?}
    % \item $L$ is the set of paths inside the infrastructure
    \item $L_j$ is the path leading the attacker to node $j$.
    \item $p$ is the probability of one monitoring node to detect an attack.
    \item  $\Pi(j)=1 - (1-p)^{|L_j\cap Mo|}$ is the probability of at least one node detecting an attack on node $j$.
    \item $\alpha$ is the probability of attacker launching an attack when transitioning to another state.
    \item $q(j)$ is the probability of node $j$ being the target of an attack
    \item $V = \{V_1,..,V_n\}, \forall i~V_i \in \mathbb{R}$ is the set containing the intrinsic value of each node.
\end{itemize}