We have implemented several use cases to determine the optimal monitoring deployment.
We propose to evaluate the MDP using three different physical topologies shown in Figure~\ref{fig:mdp-usecase-single}. The first topology is designed to be relatively balanced in terms of connections between nodes. The link between nodes 1 and 4 creates a path so each node can be reached by a single attacker in two steps.
The second topology is full-meshed and is intended to observe the monitoring strategy when the attacker only needs to compromise one node to be successful.
The last topology defines a central communication point, here node 2. This node should be of prime interest in the solution because it is on the path of most attacks in the network.

\input{figures/mdp-usecase-topo.tex}
 
We migrate a Virtual Network where the target substrate will be composed of physical nodes 1, 2 and 3.
Node 6 is the exfiltration point of the attacker, which means that the attacker will establish a path between node 6 and one of the nodes embedding the Virtual Network (\ie node 1, node 2 and/or node 3).
% In order to make the generation of the MDP states computationally tractable, we set a constant impact performance on all nodes. 
We set the discount factor of the value iteration algorithm to 0.9, a common value for use cases found in the literature.
We set $\alpha=0.7$, the probability of an attack being launched during the next transition.
We set $\beta=3$ the weight factor to target nodes supporting the embedding.
We set the intrinsic value $V$ of the nodes embedding the Virtual Network (nodes 1, 2 and 3) to be higher than the others (nodes 4, 5 and 6) because they will operate the virtual network.

We summarize these parameters in Table~\ref{tab:mdp-parameters}. 
% \CK{Precise que le choix des Vi est en accord avec le fait que les noeuds qui sont les cibles de l'embedding sont plus importants. Par ailleurs, pourquoi $c_c^j = 10$ partout ? Le cout est cense etre proportionnel a Vj non ?}
We compute the optimal policy of each MDP using MDPToolbox~\cite{Chades2014} and the Value Iteration algorithm~\cite{bellman1957}.

\begin{table}[ht]
\centering
\begin{tabular}{|c|c|l|c|c|c|c|}
\hline
$\alpha$ & \multicolumn{2}{c|}{$\beta$} & Exfiltration point & $c_f$ & $k_c$ & $\gamma$ \\ \hline
0.7      & \multicolumn{2}{c|}{3}       & Node 6             & 10    & 1    & 0.9      \\ \hline
\multicolumn{7}{|c|}{$V_1=10,V_2=10,V_3=10,V_4=5,V_5=5,V_6=5$}                          \\ \hline
\multicolumn{7}{|c|}{$c_c^1=10,c_c^2=10,c_c^3=10,c_c^4=5,c_c^5=5,c_c^6=5$}                          \\ \hline
\multicolumn{7}{|c|}{Physical embedding: Nodes 1 - Node 2 - Node 3}                     \\ \hline
\multicolumn{7}{|c|}{Migration order: 1 $\rightarrow$ 2 $\rightarrow$ 3}                \\ \hline
\end{tabular}
\caption{Use cases parameters}
\label{tab:mdp-parameters}
\end{table}



We run the MDP using different budgets and detection probabilities.
The time required to fully migrate the VN is $b_f=30$, but we explore use cases with a budget $b_f=40$ that gives more time for the attacker to compromise the infrastructure. We considered that the attacker exploits vulnerabilities that are not limited to the migration process, thus he may continue his attacks beyond the end of the migration. We discuss some implications and limitations of this assumption in Section~\ref{sec:mdp-discussion}.

The ordering of the nodes impacts the result sets of Algorithm~\ref{algo:target}, thus which nodes may be attacked at each transition.
Details about the attack source node for attacks against the migration are summarized in Table~\ref{tab:mdp-attack-source}. Each node in the \textbf{Target node} row will be attacked from the corresponding \textbf{Attack source}.

\input{figures/mdp-attack-source.tex}

\subsection{Reward of the monitoring set}
We have extracted the monitoring set of each absorbing state, and evaluated the overall reward of each monitoring set and each node individually.
We note $R(Mo)$ the reward of a monitoring set as the sum of the rewards of each corresponding absorbing state.
We also note the reward of a state (resp. a node) $R(State)$ (resp. $R(Node)$). Note however that this reward is different from the reward of an action in an MDP.

Let $S^{\text{Mo}}_{\text{abs}}$ be the set of absorbing states with a common  monitoring set $Mo$, $\rho(Mo)$ the percentage of presence of set $Mo$ in the global solution space and $R(s)$ the reward of state $s$.

\begin{equation}
    R(Mo) = \sum\limits_{s \in S^{\text{Mo}}_{\text{abs}}}\rho(Mo^s)R(s)
\end{equation}


The reward of an absorbing state $R(s)$ is the sum of the reward of each path $p$ in the MDP from the initial state and leading to the absorbing state $s$ times the percentage of presence of the $p$ in the solution space. 
Similarly, the reward of one node in the monitoring set is defined as the sum of the reward of each absorbing state where the chosen node is monitoring the infrastructure.

We illustrate these rewards in Figure~\ref{fig:absstate_reward}, where the initial state is $S_1$ and there are three absorbing states $S_4,S_5$ and $S_6$. Each state yields a reward $r_x$ with a transition probability $p_x$.
State $S_5$ is reached by two paths thus its rewards is the sum of the rewards given by each path.
% The weighted reward uses the stationary distribution of the Markov Chain corresponding to the optimal policy.
If $S_4$ and $S_6$ have the same monitoring set (\ie $Mo^{S_4} = Mo^{S_6}$), then $R(Mo^{S_4}) = R(Mo^{S_6}) = R(S_4) + R(S_6)$.
\input{figures/absorbingstate_reward.tex}

The reward of each monitoring node is computed the same way, by summing the weighted rewards of each absorbing state containing the chosen node.

For example, if $Mo^{S_4} = Mo^{S_6} =\{2,4\}$ and $Mo^{S_5} = \{1,4\}$, then the reward of node 4 is computed as $R(N_4) = R(S_4) + R(S_5) + R(S_6)$. Indeed, node 4 belongs to the monitoring set of every absorbing state.

\input{figures/mdp-usecase-results.tex}

% \FC{a relire}
\subsection{Experimental results: individual performance}
 We detail the results for each use case, and the performance of each node monitoring the infrastructure.
 Results are shown in Figure~\ref{fig:mdp-results}.
%  \GB{The figure is dense. Subfigures are too close to each other, in particular, the title of one subfigure with the caption of the below subfigure. Also captions for the right column are not centered.}.

\subsubsection*{Balanced topology: Scenarios a) and b)}
Based on the paths taken by the attacks and the ordering of the migration, we categorize the nodes into three categories: source, intermediate and border nodes. Source nodes are the origins of the attacks, intermediate nodes are the first hop from an attack node and border nodes are two hops or more away.
Scenario a) sets node 6 as the source; nodes 4 and 5 as intermediate nodes; and finally nodes 1, 2 and 3 as border nodes.
Scenario b) sets nodes 3 and 6 as sources; nodes 1 and 4 as intermediate nodes; and nodes 2 and 5 as border nodes.
The first observation is that nodes 4 and 6 are the most rewarding nodes.
This is explained as node 6 is the source of most of the attacks, and the exfiltrated data are redirected there. Node 4 is the next hop of attacks directed toward the Virtual Network (nodes 1, 2, and 3).

In both scenarios, the importance of nodes is separated according to our categorization. 
This means that the nodes pertaining to the majority of attack paths are the most rewarding.
This observation is reinforced in scenario b) where nodes 1 and 4 are closer to attack sources while nodes 2 and 5 are further away.

% In scenario a) results are stable from $p=0.6$ where node 6 overcomes node 4 in reward until $p=0.9$. All other nodes remain closely grouped, and no intermediate or border nodes is standing out.
In scenario b) the importance of node 3 increases as the importance of node 6 decreases.
The performance of the monitoring makes it more efficient to monitor the second source of attacks.


\subsubsection*{Full-mesh topology: Scenarios c) and d)}
In these scenarios, the considered topology is full-meshed, which means that each physical node is connected to every other node of the infrastructure. This topology has been chosen to minimize the impact of the routing on the detection path because the attacker is always directly connected to the target node.
Therefore, there are only two categoeries: attack nodes and intermediate nodes.
We make two observations from the results of both scenarios.
The first observation is that this topology is very sensitive to the number of attackers in the infrastructure. In the balanced topology, the nodes have the same ranking whether there is one or two attackers. This result does not hold for a full-meshed topology. In scenario c), the source of attacks globally performs better than intermediate nodes, while in scenario d) both attack and intermediate nodes have more similar rewards.

The second observation is that the detection probability tends to be more impacting in scenario d) than c). This can be explained because detection paths are more diverse with multiple attackers. Indeed, in scenario c) the focus is set on the source of attacks and then determining which nodes are the most attacked.
In scenario d) it becomes more complex to find a monitoring node that covers most attacks, as the detection can only be performed at the source of attacks and at the targeted node.

% Finally, we have examined the different optimal policies for these scenarios, and observed that some corner cases have appeared.
% Overall, the fact that for each attack only the origin of the attack and the target node may contribute to the detection has generated a diverse solution state. Simply put, 

% Each solution is composed of only two nodes. Because of the fullmesh aspect of the topology, it is complex to obtain a solution state 
% Indeed, depending on which node is targeted, there are only two nodes in the solution for the detection of the attack. and the fullmesh aspect of the topology makes it hard for those solution sets to overlap and give a better coverage.

\subsubsection*{Central topology: Scenarios e) and f)}
Using the same categorization as scenarios a) and b), we consider node 6 as a source node, node 2 as an intermediate node and nodes 1, 3, 4 and 5 as border nodes. 
Scenario f) sets nodes 1 and 6 as source nodes, nodes 2 and 4 as intermediate nodes and nodes  3 and 5 as border nodes.

In scenario e), the importance of each node reflects their role. Nodes 3, 4 and 5 do not have much impact in the solution set as they are not on the path of most attacks, while node 1 remains the main target at the beginning of the migration. The attack source, node 6, is still considered important to monitor.

Scenario f) highlights similar results. There are now two attack sources, nodes 1 and 6. Moreover, the geographical distribution of migrated nodes gives more importance to node 2 which becomes even more present on the attack paths, as shown in Table~\ref{tab:mdp-attack-source}. The fact that node 2 acts as a ``bridge" is really highlighted with two attackers.


% Optimal policies also have some corner cases with counter intuitive optimal monitoring nodes but those remain marginal considering the solution space. 

% \FC{Ajout - A relire}
\subsection{Summary}

The analysis of the results outlines several aspects common to each scenario.
The detection rate does not have a significant impact on the importance of nodes, as they remain ranked in the same order, except for scenario d) as we detailed.
% The steadiness in the evolution of each node shows that the performance of the detection is not a major factor in determining which nodes are best suited for the monitoring.
% We formulate some hypotheses in Section~\ref{sec:mdp-discussion}.

Detailed examination of the optimal policy for each budget also shows that the monitoring action is rarely used to redeploy resources elsewhere in the infrastructure.
Instead of unmonitoring nodes, the MDP chooses to do nothing to preserve the global detection probability.
Unmonitoring actions are only chosen when the the node to be unmonitored is not on the path of the next attacks, thus having the same impact as doing nothing. 
This situation happens rarely, when the attacker launches a limited number of attacks on the infrastructure.

% \input{figures/mdp-usecase-results-topofullmesh.tex}
% \input{figures/mdp-usecase-results-topo2.tex}

